---
title: "Calculate harvest rates and obtain output"
author: "Matt Russell"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
cond_true <- cond |> 
  filter(RESERVCD == 0 & SITECLCD %in% 1:6 & COND_STATUS_CD == 1 & !is.na(CONDPROP_UNADJ)) |>
  inner_join(cycles, by = "STATECD") |> 
  filter(INVYR >= recent_start & INVYR <= recent_end) |> 
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, CONDID, COND_STATUS_CD, CONDPROP_UNADJ, PROP_BASIS)
  
stand <- full_join(cond, plot, 
                   by = c("STATECD", "UNITCD", "COUNTYCD", "PLOT", "INVYR"))  |> 
  filter(RESERVCD == 0 & SITECLCD %in% 1:6 & COND_STATUS_CD == 1 & !is.na(CONDPROP_UNADJ)) |>
  inner_join(cycles, by = "STATECD") |> 
  filter(INVYR >= prev_start & INVYR <= recent_end)  
```

Filter plots to only include those that have occurred in the annual inventory 
data collection period (FIA DB Users Guide Appendix J).

```{r}
stand <- stand  |>  
  mutate(growth_cycle = ifelse(INVYR >= prev_start & INVYR <= prev_end, "PREV",
                               ifelse(INVYR >= recent_start & INVYR <= recent_end, "RECENT", NA)),
        DIST = ifelse(DSTRBCD1 > 0, "YES", "NO"),
        TRTCD = ifelse(TRTCD1 == 10 |
                         TRTCD2 == 10 |
                         TRTCD3 == 10, "TRTCD_CUT", "TRTCD_NOT_CUT"))
```

Use tree level data set to determine plot-level metrics:
Will need FIA REF_SPECIES table to calculate relative density based on 
Ducey and Knapp equation.

```{r}
tree <- tree |>  
  inner_join(cycles, by = "STATECD") |> 
  filter(INVYR >= prev_start & INVYR <= recent_end)

ref_spp <- read_csv('C:/Users/matt/Documents/Arbor/Projects/FFCP/harvesting/data/REF_SPECIES.csv') %>% 
  select(SPCD, WOOD_SPGR_MC12VOL_DRYWT)

tree <- tree |> 
  inner_join(ref_spp, by = "SPCD") |> 
  inner_join(stand, by = c("STATECD", "UNITCD", "COUNTYCD", "PLOT", "INVYR"))

```

Calculate metrics:

```{r}
# Live trees
tree_live <- tree %>% 
  filter(STATUSCD == 1) %>% 
  mutate(tpa_part = TPA_UNADJ,
         tpa_part_5 = ifelse(DIA >= 5.0, TPA_UNADJ, 0),
         ba_part = (0.00545415*DIA^2)*TPA_UNADJ,
         ba_part_5 = ifelse(DIA >= 5.0, (0.00545415*DIA^2)*TPA_UNADJ, 0),
         RD_part = ifelse(DIA >= 5.0,
                          ((0.00015+(0.00218*WOOD_SPGR_MC12VOL_DRYWT))*
                             (DIA*2.54/10)**1.6)*2.4701, 0))

tree_live_sum <- tree_live %>% 
  group_by(STATECD, UNITCD, COUNTYCD, PLOT, INVYR) %>% 
  summarize(TPA = sum(tpa_part, na.rm = T),
            BA = sum(ba_part, na.rm = T),
            TPA_5 = sum(tpa_part_5, na.rm = T),
            BA_5 = sum(ba_part_5, na.rm = T),
            RD = sum(RD_part, na.rm = T)) %>% 
  mutate(QMD = sqrt((BA_5/TPA_5)/(0.00545415))) %>%
  mutate_at(vars(TPA, BA, RD, QMD), replace_na, 0)

# Removed trees
tree_removed <- tree %>% 
  filter(STATUSCD == 3 & PREV_STATUS_CD == 1) %>% 
  mutate(TPAREMV_UNADJ = ifelse(PREVDIA >= 5.0, 6.018046, 74.965282),
         tpa_rem_part = TPAREMV_UNADJ,
         tpa_rem_part_5 = ifelse(PREVDIA >= 5.0, TPAREMV_UNADJ, 0),
         ba_rem_part = (0.00545415*PREVDIA^2)*TPAREMV_UNADJ,
         ba_rem_part_5 = ifelse(PREVDIA >= 5.0, (0.00545415*PREVDIA^2)*TPAREMV_UNADJ, 0),
         RD_rem_part = ifelse(PREVDIA >= 5.0,
                          ((0.00015+(0.00218*WOOD_SPGR_MC12VOL_DRYWT))*
                             (PREVDIA*2.54/10)**1.6)*2.4701, 0))

tree_removed_sum <- tree_removed %>% 
  group_by(STATECD, UNITCD, COUNTYCD, PLOT, INVYR) %>% 
  summarize(TPA_rem = sum(tpa_rem_part),
            BA_rem = sum(ba_rem_part),
            TPA_rem_5 = sum(tpa_rem_part_5),
            BA_rem_5 = sum(ba_rem_part_5),
            RD_rem = sum(RD_rem_part)) %>% 
  mutate(QMD_rem = sqrt((BA_rem_5/TPA_rem_5)/(0.00545415))) %>%
  mutate_at(vars(TPA_rem, BA_rem, RD_rem, QMD_rem), replace_na, 0)

tree_sum <- left_join(tree_live_sum, tree_removed_sum, 
                       by = c("STATECD", "UNITCD", "COUNTYCD", "PLOT", "INVYR"))

stand <- inner_join(stand, tree_sum, by = c("STATECD", "UNITCD", "COUNTYCD", "PLOT", "INVYR")) %>% 
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, RESERVCD, OWNCD, OWNGRPCD, COND_STATUS_CD ,
         FORINDCD, ADFORCD, FORTYPCD, STDAGE, STDSZCD, SITECLCD, SICOND, SIBASE, SISP,
         STDORGCD, STDORGSP, SLOPE, ASPECT, PHYSCLCD, GSSTKCD, ALSTKCD, DSTRBCD1, 
         DSTRBYR1, DSTRBCD2, DSTRBYR2, DSTRBCD3, DSTRBYR3, TRTCD, TRTCD1, TRTYR1, TRTCD2,
         TRTYR2, TRTCD3, TRTYR3, WATERCD, LAT, LON, growth_cycle, CONDPROP_UNADJ, PROP_BASIS,
         ELEV, ECOSUBCD, DIST, RDDISTCD, CONDID,
         TPA, BA, RD, QMD,
         TPA_rem, BA_rem, BA_rem_5, RD_rem, QMD_rem)
```

Create time1, time 2 data set. Variables with `_PREV` designation indicate 
measurement at time 1:

```{r}
stand_prev <- stand %>%
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) %>%
  mutate(INVYR_PREV = lag(INVYR),
         TRTCD_PREV = lag(TRTCD),
         GSSTKCD_PREV = lag(GSSTKCD),
         STDAGE_PREV = lag(STDAGE),
         RESERVCD_PREV = lag(RESERVCD),
         FORINDCD_PREV = lag(FORINDCD),
         SITECLCD_PREV = lag(SITECLCD),
         FORTYPCD_PREV = lag(FORTYPCD),
         STDSZCD_PREV = lag(STDSZCD),
         OWNCD_PREV = lag(OWNCD),
         OWNGRPCD_PREV = lag(OWNGRPCD),
         STDORGCD_PREV = lag(STDORGCD),
         SICOND_PREV = lag(SICOND),
         BA_PREV = lag(BA),
         TPA_PREV = lag(TPA), 
         RD_PREV = lag(RD),
         QMD_PREV = lag(QMD),
         TPA_rem_PREV = lag(TPA_rem), 
         BA_rem_PREV = lag(BA_rem),
         BA_rem_5_PREV = lag(BA_rem_5), 
         RD_rem_PREV = lag(RD_rem), 
         QMD_rem_PREV = lag(QMD_rem)) %>% 
  
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, INVYR_PREV, TRTCD_PREV,
         GSSTKCD_PREV, STDAGE_PREV, RESERVCD_PREV, FORINDCD_PREV, SITECLCD_PREV,
         SICOND_PREV, FORTYPCD_PREV, STDSZCD_PREV, OWNCD_PREV, COND_STATUS_CD ,
         OWNGRPCD_PREV, STDORGCD_PREV, INVYR_PREV, BA_PREV, TPA_PREV, RD_PREV, 
         QMD_PREV, CONDPROP_UNADJ, PROP_BASIS, CONDID,
         TPA_rem_PREV, BA_rem_PREV, BA_rem_5_PREV, RD_rem_PREV, QMD_rem_PREV) %>% 
  
  arrange(STATECD, UNITCD, COUNTYCD, PLOT, INVYR)

stand <- inner_join(stand, stand_prev) 
```

Merge stand file to forest type and forest type group tables, obtained from
the FIA reference tables:

```{r}
REF_FOREST_TYPE <- read_csv('C:/Users/matt/Documents/Arbor/Projects/FFCP/harvesting/data/REF_FOREST_TYPE.csv')
REF_FOREST_TYPE_GROUP <- read_csv('C:/Users/matt/Documents/Arbor/Projects/FFCP/harvesting/data/REF_FOREST_TYPE_GROUP.csv')

forest_type <- REF_FOREST_TYPE %>% 
  rename(FORTYPCD_PREV = VALUE,
         FORTYPCD_NAME_PREV = MEANING,
         FORTYPGRPCD_PREV = TYPGRPCD) %>% 
  select(FORTYPCD_PREV, FORTYPCD_NAME_PREV, FORTYPGRPCD_PREV)

forest_group <- REF_FOREST_TYPE_GROUP %>% 
  rename(FORTYPGRPCD_NAME_PREV = MEANING,
         FORTYPGRPCD_PREV = VALUE) %>% 
    select(FORTYPGRPCD_PREV, FORTYPGRPCD_NAME_PREV)

forest <- inner_join(forest_type, forest_group, by = "FORTYPGRPCD_PREV")

stand <- inner_join(stand, forest, by = "FORTYPCD_PREV") 
```

# Determine harvest amounts
```{r}
stand <- stand |> 
  filter(growth_cycle %in% c("RECENT")) |> 
  mutate(# Approach 1 (TRTCD): Treatment code (TRTCD):
         BA_change = ((BA_PREV - BA)/BA_PREV),
         CUTTING_TRTCD_MAIN = ifelse(TRTCD == "TRTCD_CUT", "YES", "NO"),
         TRTCD_new = ifelse(CUTTING_TRTCD_MAIN %in% c("YES") & 
                                      DIST %in% c("NO"), 
                                1, 0),
         # Approach 2 (BA_RED_25): Basal area, stand level > 25% (CUTTING_BA_25_STAND):
         CUTTING_BA_25_MAIN = ifelse(BA_change >= 0.25, "YES", "NO"),
         BA_RED = ifelse(CUTTING_BA_25_MAIN %in% c("YES") & 
                                      DIST %in% c("NO"), 1, 0),
         # Approach 3 (BA_RED_TREE_25): Basal area, tree level (CUTTING_BA_TREE):
         BA_change_TREE = (BA_rem + BA) - BA,
         CUTTING_BA_TREE_MAIN = ifelse(BA_change_TREE >= 0.25, "YES", "NO"),
         BA_RED_TREE = ifelse(CUTTING_BA_TREE_MAIN %in% c("YES") & 
                                      DIST %in% c("NO"), 1, 0),
         # Approach 4 (RD_RED_10): Relative density, stand level (CUTTING_RD_STAND):
         RD_change = RD_PREV - RD,
         CUTTING_RD_MAIN = ifelse(RD_change >= 0.25, "YES", "NO"),
         RD_RED = ifelse(CUTTING_RD_MAIN %in% c("YES") & 
                                      DIST %in% c("NO"), 1, 0),
         
         # Approach 5 (RD_RED_TREE_10): Relative density, tree level (CUTTING_RD_TREE):
         RD_change_TREE = (RD_rem + RD) - RD,
         CUTTING_RD_TREE_MAIN = ifelse(RD_change_TREE >= 0.25, "YES", "NO"),
         RD_RED_TREE = ifelse(CUTTING_RD_TREE_MAIN %in% c("YES") & 
                                      DIST %in% c("NO"), 1, 0),
         # Approach 6 (TRTCD + BA_RED_25): Basal area removed > 25% (CUTTING_BA_25_STAND) OR
         TRTCD_BA_RED = ifelse(BA_RED == 1|
                                              (TRTCD_new == 1 & BA_change > 0),
                                            1, 0),
         # Approach 7 (TRTCD + BA_RED_TREE): 
         TRTCD_BA_RED_TREE = ifelse(BA_RED_TREE == 1 |
                                              (TRTCD_new == 1 & BA_change > 0),
                                            1, 0),
         # Approach 8 (TRTCD + RD_RED): Basal area removed > 25% (CUTTING_RD_25_STAND) OR
         TRTCD_RD_RED = ifelse(RD_RED == 1 |
                                             (TRTCD_new == 1 & BA_change > 0),
                                            1, 0),
         # Approach 9 (TRTCD + RD_RED_TREE): 
         TRTCD_RD_RED_TREE = ifelse(RD_RED_TREE == 1 |
                                             (TRTCD_new == 1 & BA_change > 0),
                                            1, 0)
         ) |> 
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, CONDID, 
         TRTCD_new, BA_RED, BA_RED_TREE, RD_RED, RD_RED_TREE,
         TRTCD_BA_RED, TRTCD_BA_RED_TREE, TRTCD_RD_RED, TRTCD_RD_RED_TREE)


cond_true2 <- cond_true |> 
  left_join(stand, by = c("STATECD", "UNITCD","COUNTYCD", "PLOT", "INVYR", "CONDID")) 
```

```{r}
write_csv(cond_true2, paste0("C:/Users/matt/Documents/Arbor/Projects/FFCP/harvesting/data/v2/", 
                             state_abbr, 
                             "_COND_HARVEST_MIDPT.csv"))
```

